Proofs for ``Geometry of Local Spectral Clustering''

### 2021-01-13 jmlr/revision/proofs.tex

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\SM}{\mathrm{SM}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}

@@

The proofs of these upper bounds will be similar to each other (in places word-for-word alike), and will follow a similar approach and use similar notation to that of \citep{lovasz1990,andersen2006}. For $h: [0,2m] \to [0,1]$, $0 \leq K_0 \leq m$ and $k \in [K_0,2m - K_0]$, define
\begin{equation*}
L_{K_0}(k;h) = \frac{2m - K_0 - k}{2m - 2K_0}h(K_0) + \frac{k - K_0}{2m - 2K_0}h(2m - K_0)
\end{equation*}
to be the linear interpolant of $h(K_0)$ and $h(2m - K_0)$, and additionally let
\begin{equation*}
C(K_0;h) := \max\set{\frac{h(k) - L_{K_0}(k;h)}{\sqrt{\overline{k}}}: K_0 \leq k \leq  2m - K_0}.
\end{equation*}
where we use the notation $\overline{k} := \min\{k, 2m - k\}$, and treat $0/0$ as equal to $1$. Our first pair of Lemmas upper bound $h^{(t)}$ and $h^{(\alpha)}$ as a function of $L_{K_0}$ and $C(K_0)$. Lemma~\ref{lem:mixing_random_walk} implies that if $t$ is large relative to $\Psi(G)$, then $h^{(t)}(\cdot)$ must be small.

@@

\begin{lemma}[\textbf{c.f Theorem~3 of \citep{andersen2006}}]
	\label{lem:mixing_time_PPR}
	Let $\phi \in [0,1]$. Either the following bound holds for any $t \in \mathbb{N}$, any $K_0 \in [0,m]$, and any $k \in [K_0,2m - K_0]$:
	\begin{equation}
	\label{eqn:mixing_time_PPR}
	h^{(\alpha)}(k) \leq \alpha t + L_{K_0}(k; h^{(\alpha)}) + C(K_0;h^{(\alpha)})\sqrt{\overline{k}}\left(1 - \frac{\phi^2}{8}\right)^t
	\end{equation}
	or else there exists some sweep cut $S_{j}$ of $p_v$ such that $\Phi(S_j;G) < \phi$.
\end{lemma}

@@

We next establish an upper bound on $C_{K_0}(k;h)$, which rests on the following key observation: since $h(k)$ is concave and $L_{K_0}(K_0;h) = h(K_0)$, it holds that
\begin{equation}
\label{pf:linearization_bound_1}
\frac{h(k) - L_{K_0}(k)}{\sqrt{\overline{k}}} \leq
\begin{cases}
h'(K_0) \sqrt{k},~& k \leq m \\
-h'(2m - K_0) \sqrt{2m - k},~& k > m.
\end{cases}
\end{equation}
(Since $h$ is not differentiable at $k = k_j$, here $h'$ refers to the right derivative of $h$.)

@@

	\textbf{Case 1.}
	Assume $k_j - \Psi(G) \overline{k}_j$ and $k_j + 2 \Psi(G) \overline{k}_j$ are both in $[K_0,2m  - K_0]$. We are therefore in a position to apply our inductive hypothesis to both terms on the right hand side of~\eqref{pf:mixing_random_walk_inductive_step_h} and obtain the following:
	\begin{align*}
	h^{(t)}(k_j) & \leq \frac{1}{2}\biggl(L_{K_0}\bigl(k_j - \Psi(G) \overline{k}_j; h^{(t_0)}\bigr) + L_{K_0}\bigl(k_j + \Psi(G) \overline{k}_j; h^{(t_0)}\bigr) \biggr) ~~+ \\
	& \quad~ \frac{1}{2}C\bigl(K_0; h^{(t_0)}\bigr) \cdot \Bigl(\sqrt{\overline{k_j - \Psi(G) \overline{k}_j}} + \sqrt{\overline{k_j + \Psi(G) \overline{k}_j}}\Bigr)\left(1 - \frac{\Psi(G)^2}{8}\right)^{t-t_0 - 1} \\
	& = L_{K_0}(k;h^{(t_0)}) + \frac{1}{2}\biggl(C(K_0;h^{t_0})\bigl(\sqrt{\overline{k_j - \Psi(G) \overline{k}_j}} + \sqrt{\overline{k_j + \Psi(G) \overline{k}_j}}\bigr)\left(1 - \frac{\Psi(G)^2}{8}\right)^{t-t_0 - 1} \biggr) \\
	& \leq L_{K_0}(k;h^{(t_0)}) + \frac{1}{2}\biggl(C(K_0;h^{(t_0)})\bigl(\sqrt{\overline{k}_j - \Psi(G) \overline{k}_j} + \sqrt{\overline{k}_j + \Psi(G) \overline{k}_j}\bigr)\left(1 - \frac{\Psi(G)^2}{8}\right)^{t-t_0 - 1} \biggr).
	\end{align*}
	A Taylor expansion of $\sqrt{1 + \Psi(G)}$ around $\Psi(G) = 0$ yields the following bound:
	\begin{equation*}
	\sqrt{1 + \Psi(G)} + \sqrt{1 - \Psi(G)} \leq 2 - \frac{\Psi(G)^2}{4},
	\end{equation*}
	and therefore
	\begin{align*}
	h^{(t)}(k_j) & \leq L_{K_0}(k;h^{(t_0)}) + \frac{C(K_0;h^{(t_0)})}{2}\cdot \sqrt{\overline{k}_j}\cdot\left(2 - \frac{\Psi(G)^2}{4}\right)\left(1 - \frac{\Psi(G)^2}{8}\right)^{t-1} \\
	&= L_{K_0}(k_j;h^{(t_0)}) + C(K_0;h^{(t_0)})\sqrt{\overline{k}_j}\left(1 - \frac{\Psi(G)^2}{8}\right)^{t - t_0}.
	\end{align*}

	\textbf{Case 2.} Otherwise one of $k_j - 2 \Psi(G) \overline{k}_j$ or $k_j + 2 \Psi(G) \overline{k}_j$ is not in $[K_0,2m  - K_0]$. Without loss of generality assume $k_j < m$, so that (i) we have $k_j - 2 \Psi(G) \overline{k}_j < K_0$ and (ii) $k_j + (k_j - K_0) \leq 2m - K_0$. We deduce the following:
	\begin{align*}
	h^{(t)}(k_j) & \overset{(\textrm{i})}{\leq} \frac{1}{2}\Bigl(h^{(t - 1)}(K_0) + h^{(t - 1)}\bigl(k_j + (k_j - K_0)\bigr)\Bigr) \\
	& \overset{(\textrm{ii})}{\leq} \frac{1}{2}\Bigl(h^{(t_0)}(K_0) + h^{(t)}\bigl(k_j + (k_j - K_0)\bigr)\Bigr) \\
	& \overset{(\textrm{iii})}{\leq}\frac{1}{2}\Bigl(L_{K_0}(K_0;h^{(t_0)}) + L_{K_0}(2k_j - K_0; h^{(t_0)}\bigr) + C(K_0;h^{(t_0)})\sqrt{\overline{2k_j - K_0}}\left(1 - \frac{\Psi(G)^2}{8}\right)^{t - t_0 - 1}\Bigr) \\
	& \leq L_{K_0}(k_j; h^{(t_0)}) + C(K_0;h^{(t_0)}) \frac{\sqrt{2\overline{k}_j}}{2} \left(1 - \frac{\Psi(G)^2}{8}\right)^{t - t_0 - 1} \\
	& \leq L_{K_0}(k_j;h^{(t_0)}) + C(K_0;h^{(t_0)}) \sqrt{\overline{k}_j} \cdot \left(1 - \frac{\Psi(G)^2}{8}\right)^{t - t_0}
	\end{align*}
	where $(\textrm{(i)})$ follows from~\eqref{pf:mixing_random_walk_inductive_step_h} and the concavity of $h^{(t - 1)}$,  we deduce $(\textrm{ii})$ from~\eqref{pf:mixing_random_walk_inductive_step_h}, which implies that $h^{(t)}(k) \leq h^{(t_0)}(k)$, and $(\textrm{iii})$ follows from applying the inductive hypothesis to $h^{(t - 1)}(2k_j - K_0)$. 	
\end{proof}

@@

	We proceed by induction on $t$. Our base case will be $t = 0$. Observe that $C(K_0;h^{(\alpha)}) \cdot \sqrt{\overline{k}} \geq h^{(\alpha)}(k) - L_{K_0}(k;h^{(\alpha)})$ for all $k \in [K_0,2m - K_0]$, which implies
	\begin{equation*}
	L_{K_0}(k;h^{(\alpha)}) + C(K_0;h^{(\alpha)}) \cdot \sqrt{\overline{k}} \geq h^{(\alpha)}(k).
	\end{equation*}

@@

	By Lemma 5 of \citet{andersen2006}, we have that
	\begin{align}
	p_v[k_j] & \leq \alpha + \frac{1}{2} \bigl(p_v[k_j - \mathrm{cut}(S_j;G)] + p_v[k_j + \mathrm{cut}(S_j;G)] \bigr) \nonumber\\
	& \leq \alpha + \frac{1}{2} \bigl(p_v[k_j - \Phi(S_j;G) \overline{k}_j] + p_v[k_j + \Phi(S_j;G) \overline{k}_j]  \bigr) \nonumber \\
	& \leq \alpha + \frac{1}{2} \bigl(p_v[k_j - \phi \overline{k}_j] + p_v[k_j + \phi \overline{k}_j]\bigr) \nonumber
	\end{align}
	and subtracting $k_j/2m$ from both sides, we get
	\begin{equation}
	\label{eqn:mixing_time_PPR_pf1}
	h^{(\alpha)}(k_j) \leq \alpha + \frac{1}{2} \bigl(h^{(\alpha)}(k_j - \phi \overline{k}_j) + h^{(\alpha)}(k_j +  \phi \overline{k}_j) \bigr)
	\end{equation}
	From this point, we divide our analysis into cases.

	\textbf{Case 1.}
	Assume $k_j - 2 \phi \overline{k}_j$ and $k_j + 2 \phi \overline{k}_j$ are both in $[K_0,2m  - K_0]$. We are therefore in a position to apply our inductive hypothesis to \eqref{eqn:mixing_time_PPR_pf1}, yielding
	\begin{align*}
	h^{(\alpha)}(k_j) & \leq \alpha + \alpha(t-1) \frac{1}{2}\biggl(L_{K_0}(k_j - \phi \overline{k}_j) + L_{K_0}(k_j + \phi \overline{k}_j) + C(K_0;h^{(\alpha)})\bigl(\sqrt{\overline{k_j - \phi \overline{k}_j}} + \sqrt{\overline{k_j + \phi \overline{k}_j}}\bigr)\left(1 - \frac{\phi^2}{8}\right)^{t-1} \biggr) \\
	& \leq \alpha t + L_{K_0}(k;h^{(\alpha)}) + \frac{1}{2}\biggl(C(K_0;h^{(\alpha)})\bigl(\sqrt{\overline{k_j - \phi \overline{k}_j}} + \sqrt{\overline{k_j + \phi \overline{k}_j}}\bigr)\left(1 - \frac{\phi^2}{8}\right)^{t-1} \biggr) \\
	& \leq \alpha t + L_{K_0}(k;h^{(\alpha)}) + \frac{1}{2}\biggl(C(K_0;h^{(\alpha)})\bigl(\sqrt{\overline{k}_j - \phi \overline{k}_j} + \sqrt{\overline{k}_j + \phi \overline{k}_j}\bigr)\left(1 - \frac{\phi^2}{8}\right)^{t-1} \biggr).
	\end{align*}
	and therefore
	\begin{equation*}
	h^{(\alpha)}(k_j) \leq  \alpha t + L_{K_0}(k;h^{(\alpha)}) + \frac{C(K_0;h^{(\alpha)})}{2}\cdot \sqrt{\overline{k}_j}\cdot\left(2 - \frac{\phi^2}{4}\right)\left(1 - \frac{\phi^2}{8}\right)^{t-1} = \alpha t + L_{K_0}(k;h^{(\alpha)}) + C(K_0;h^{(\alpha)})\sqrt{\overline{k}_j}\left(1 - \frac{\phi^2}{8}\right)^{t}.
	\end{equation*}

	\textbf{Case 2.} Otherwise one of $k_j - 2 \phi \overline{k}_j$ or $k_j + 2 \phi \overline{k}_j$ is not in $[K_0,2m  - K_0]$. Without loss of generality assume $k_j < m$, so that (i) we have $k_j - 2 \phi \overline{k}_j < K_0$ and (ii) $k_j + (k_j - K_0) \leq 2m - K_0$. By the concavity of $h$, and applying the inductive hypothesis to $h^{(\alpha)}2k_j - K_0)$, we have
	\begin{align*}
	h^{(\alpha)}(k_j) & \leq \alpha + \frac{1}{2}\Bigl(h^{(\alpha)}(K_0) + h\bigl(k_j + (k_j - K_0)\bigr)\Bigr) \\
	& \leq\alpha + \frac{\alpha(t - 1)}{2} + \frac{1}{2}\Bigl(L_{K_0}(K_0;p^{\alpha}) + L_{K_0}(2k_j - K_0\bigr) + C(K_0;h^{(\alpha)})\sqrt{\overline{2k_j - K_0}}\left(1 - \frac{\phi^2}{8}\right)^{t - 1}\Bigr) \\
	& \leq \alpha t + L_{K_0}(k_j) + C(K_0;h^{(\alpha)}) \frac{\sqrt{2\overline{k}_j}}{2} \left(1 - \frac{\phi^2}{8}\right)^{t - 1} \\
	& \leq \alpha t + L_{K_0}(k_j) + C(K_0;h^{(\alpha)}) \sqrt{\overline{k}_j} \cdot \left(1 - \frac{\phi^2}{8}\right)^{t}
	\end{align*}
\end{proof}

@@

\section{Sample-to-population bounds}
\label{sec:sample_to_population}
In this section, we prove Proposition~\ref{prop:sample_to_population_1}, by establishing high-probability finite-sample bounds on various functionals of the random graph $G_{n,r}$: cut, volume, and normalized cut~(\ref{subsec:sample_to_population_ncut}), minimum and maximum degree, and local spread~(\ref{subsec:sample_to_population_local_spread}), and conductance~(\ref{subsec:sample_to_population_conductance}). To establish these results, we will use several different concentration inequalities, and we begin by reviewing these in~(\ref{subsec:concentration}). Throughout, we denote the empirical probability of a set $\mc{S} \subseteq \Rd$ as $\Pbb_n(\mc{S}) = \sum_{i = 1}^{n} \1\{x_i \in \mc{S}\}/n$, and the conditional (on being in $\mc{C}$) empirical probability as $\wt{\Pbb}_n = \sum_{i = 1}^{n} \1\{x_i \in (\mc{S} \cap \mc{C})\}/\wt{n}$, where $\wt{n} = |\mc{C}[X]|$ is the number of sample points that are in $\mc{C}$. For a probability measure $\Qbb$, we also write
\begin{equation*}
d_{\min}(\Qbb) := \inf_{x \in \mathrm{supp}(\mbb{Q})} \deg_{\Pbb,r}(x),~~\textrm{and}~~d_{\max}(\Qbb) := \sup_{x \in \mathrm{supp}(\mbb{Q})} \deg_{\Pbb,r}(x).
\end{equation*}

@@

Many graph functionals are order-2 U-statistics, and we use Hoeffding's inequality to control the deviations of these functionals from their expectations. Recall that $U_n$ is an order-2 U-statistic with kernel $\varphi: \Rd \times \Rd \to \Reals$ if 
\begin{equation*}
U_n = \frac{1}{n(n-1)}\sum_{i = 1}^{n} \sum_{j \neq i} \varphi(x_i,x_j).
\end{equation*}
We write $\norm{\varphi}_{\infty} = \sup_{x,y} |\varphi(x,y)|$. 
\begin{lemma}[Hoeffding's Inequality for U-statistics.]
	\label{lem:hoeffding}
	Fix $\delta \in (0,1)$. Assume $\norm{\varphi}_{\infty} \leq 1$. Then,
	\begin{equation*}
	(1 - \delta) \mathbb{E}[U_n] \leq U_n \leq (1 + \delta) \mathbb{E}[U_n]
	\end{equation*}
	with probability at least $1 - 2 \exp\left(-\delta^2 (\mathbb{E}[U_n])^2 n\right)$.
\end{lemma}

@@

	For any $x \in \Rd$, the remainder terms can be upper bounded: for any $\Delta_{12} \geq 0$,

@@

	and for any $0 \leq \Delta_{12} \leq r/(4(d - 1))$,

@@

	To derive~\eqref{pf:conductance_lb_transportation_distance_3}, suppose that $\mc{R} \subseteq \mathrm{supp}(\mbb{Q}_2)$ and $\mc{R} \not\in \mathfrak{L}(\mbb{Q}_1,\mbb{Q}_2)$. Without loss of generality, we shall assume that $\vol_{\mbb{Q}_2,r}(\mc{R}) \leq (1 - \Psi_r(\mbb{Q}_1)) \cdot d_{\min}(\mbb{Q}_2)^2$ (otherwise we can work with respect to $\mc{R}^c$.)  Then, for all $x \in \mc{R}$:

@@

\section{aPPR and Misclassification Error}
\label{sec:appr_misclassification_error}

>>>>

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\wb{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\SM}{\mathrm{SM}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}

@@

The proofs of these upper bounds will be similar to each other (in places word-for-word alike), and will follow a similar approach and use similar notation to that of \citep{lovasz1990,andersen2006}. For $h: [0,2m] \to [0,1]$, $0 \leq K_0 \leq m$ and $k \in [K_0,2m - K_0]$, define
\begin{equation*}
L_{K_0}(k;h) = \frac{2m - K_0 - k}{2m - 2K_0}h(K_0) + \frac{k - K_0}{2m - 2K_0}h(2m - K_0)
\end{equation*}
to be the linear interpolant of $h(K_0)$ and $h(2m - K_0)$, and additionally let
\begin{equation*}
C(K_0;h) := \max\set{\frac{h(k) - L_{K_0}(k;h)}{\sqrt{\wb{k}}}: K_0 \leq k \leq  2m - K_0}.
\end{equation*}
where we use the notation $\wb{k} := \min\{k, 2m - k\}$, and treat $0/0$ as equal to $1$. Our first pair of Lemmas upper bound $h^{(t)}$ and $h^{(\alpha)}$ as a function of $L_{K_0}$ and $C(K_0)$. Lemma~\ref{lem:mixing_random_walk} implies that if $t$ is large relative to $\Psi(G)$, then $h^{(t)}(\cdot)$ must be small.

@@

\begin{lemma}[\textbf{c.f Theorem~3 of \citep{andersen2006}}]
	\label{lem:mixing_time_PPR}
	Let $\phi \in [0,1]$. Either the following bound holds for any $t \in \mathbb{N}$, any $K_0 \in [0,m]$, and any $k \in [K_0,2m - K_0]$:
	\begin{equation}
	\label{eqn:mixing_time_PPR}
	h^{(\alpha)}(k) \leq \alpha t + L_{K_0}(k; h^{(\alpha)}) + C(K_0;h^{(\alpha)})\sqrt{\wb{k}}\left(1 - \frac{\phi^2}{8}\right)^t
	\end{equation}
	or else there exists some sweep cut $S_{j}$ of $p_v$ such that $\Phi(S_j;G) < \phi$.
\end{lemma}

@@

We next establish an upper bound on $C_{K_0}(k;h)$, which rests on the following key observation: since $h(k)$ is concave and $L_{K_0}(K_0;h) = h(K_0)$, it holds that
\begin{equation}
\label{pf:linearization_bound_1}
\frac{h(k) - L_{K_0}(k)}{\sqrt{\wb{k}}} \leq
\begin{cases}
h'(K_0) \sqrt{k},~& k \leq m \\
-h'(2m - K_0) \sqrt{2m - k},~& k > m.
\end{cases}
\end{equation}
(Since $h$ is not differentiable at $k = k_j$, here $h'$ refers to the right derivative of $h$.)

@@

	\textbf{Case 1.}
	Assume $k_j - \Psi(G) \wb{k}_j$ and $k_j + 2 \Psi(G) \wb{k}_j$ are both in $[K_0,2m  - K_0]$. We are therefore in a position to apply our inductive hypothesis to both terms on the right hand side of~\eqref{pf:mixing_random_walk_inductive_step_h} and obtain the following:
	\begin{align*}
	h^{(t)}(k_j) & \leq \frac{1}{2}\biggl(L_{K_0}\bigl(k_j - \Psi(G) \wb{k}_j; h^{(t_0)}\bigr) + L_{K_0}\bigl(k_j + \Psi(G) \wb{k}_j; h^{(t_0)}\bigr) \biggr) ~~+ \\
	& \quad~ \frac{1}{2}C\bigl(K_0; h^{(t_0)}\bigr) \cdot \Bigl(\sqrt{\wb{k_j - \Psi(G) \wb{k}_j}} + \sqrt{\wb{k_j + \Psi(G) \wb{k}_j}}\Bigr)\left(1 - \frac{\Psi(G)^2}{8}\right)^{t-t_0 - 1} \\
	& = L_{K_0}(k;h^{(t_0)}) + \frac{1}{2}\biggl(C(K_0;h^{t_0})\bigl(\sqrt{\wb{k_j - \Psi(G) \wb{k}_j}} + \sqrt{\wb{k_j + \Psi(G) \wb{k}_j}}\bigr)\left(1 - \frac{\Psi(G)^2}{8}\right)^{t-t_0 - 1} \biggr) \\
	& \leq L_{K_0}(k;h^{(t_0)}) + \frac{1}{2}\biggl(C(K_0;h^{(t_0)})\bigl(\sqrt{\wb{k}_j - \Psi(G) \wb{k}_j} + \sqrt{\wb{k}_j + \Psi(G) \wb{k}_j}\bigr)\left(1 - \frac{\Psi(G)^2}{8}\right)^{t-t_0 - 1} \biggr).
	\end{align*}
	A Taylor expansion of $\sqrt{1 + \Psi(G)}$ around $\Psi(G) = 0$ yields the following bound:
	\begin{equation*}
	\sqrt{1 + \Psi(G)} + \sqrt{1 - \Psi(G)} \leq 2 - \frac{\Psi(G)^2}{4},
	\end{equation*}
	and therefore
	\begin{align*}
	h^{(t)}(k_j) & \leq L_{K_0}(k;h^{(t_0)}) + \frac{C(K_0;h^{(t_0)})}{2}\cdot \sqrt{\wb{k}_j}\cdot\left(2 - \frac{\Psi(G)^2}{4}\right)\left(1 - \frac{\Psi(G)^2}{8}\right)^{t-1} \\
	&= L_{K_0}(k_j;h^{(t_0)}) + C(K_0;h^{(t_0)})\sqrt{\wb{k}_j}\left(1 - \frac{\Psi(G)^2}{8}\right)^{t - t_0}.
	\end{align*}

	\textbf{Case 2.} Otherwise one of $k_j - 2 \Psi(G) \wb{k}_j$ or $k_j + 2 \Psi(G) \wb{k}_j$ is not in $[K_0,2m  - K_0]$. Without loss of generality assume $k_j < m$, so that (i) we have $k_j - 2 \Psi(G) \wb{k}_j < K_0$ and (ii) $k_j + (k_j - K_0) \leq 2m - K_0$. We deduce the following:
	\begin{align*}
	h^{(t)}(k_j) & \overset{(\textrm{i})}{\leq} \frac{1}{2}\Bigl(h^{(t - 1)}(K_0) + h^{(t - 1)}\bigl(k_j + (k_j - K_0)\bigr)\Bigr) \\
	& \overset{(\textrm{ii})}{\leq} \frac{1}{2}\Bigl(h^{(t_0)}(K_0) + h^{(t)}\bigl(k_j + (k_j - K_0)\bigr)\Bigr) \\
	& \overset{(\textrm{iii})}{\leq}\frac{1}{2}\Bigl(L_{K_0}(K_0;h^{(t_0)}) + L_{K_0}(2k_j - K_0; h^{(t_0)}\bigr) + C(K_0;h^{(t_0)})\sqrt{\wb{2k_j - K_0}}\left(1 - \frac{\Psi(G)^2}{8}\right)^{t - t_0 - 1}\Bigr) \\
	& \leq L_{K_0}(k_j; h^{(t_0)}) + C(K_0;h^{(t_0)}) \frac{\sqrt{2\wb{k}_j}}{2} \left(1 - \frac{\Psi(G)^2}{8}\right)^{t - t_0 - 1} \\
	& \leq L_{K_0}(k_j;h^{(t_0)}) + C(K_0;h^{(t_0)}) \sqrt{\wb{k}_j} \cdot \left(1 - \frac{\Psi(G)^2}{8}\right)^{t - t_0}
	\end{align*}
	where $(\textrm{(i)})$ follows from~\eqref{pf:mixing_random_walk_inductive_step_h} and the concavity of $h^{(t - 1)}$,  we deduce $(\textrm{ii})$ from~\eqref{pf:mixing_random_walk_inductive_step_h}, which implies that $h^{(t)}(k) \leq h^{(t_0)}(k)$, and $(\textrm{iii})$ follows from applying the inductive hypothesis to $h^{(t - 1)}(2k_j - K_0)$. 	
\end{proof}

@@

	We proceed by induction on $t$. Our base case will be $t = 0$. Observe that $C(K_0;h^{(\alpha)}) \cdot \sqrt{\wb{k}} \geq h^{(\alpha)}(k) - L_{K_0}(k;h^{(\alpha)})$ for all $k \in [K_0,2m - K_0]$, which implies
	\begin{equation*}
	L_{K_0}(k;h^{(\alpha)}) + C(K_0;h^{(\alpha)}) \cdot \sqrt{\wb{k}} \geq h^{(\alpha)}(k).
	\end{equation*}

@@

	By Lemma 5 of \citet{andersen2006}, we have that
	\begin{align}
	p_v[k_j] & \leq \alpha + \frac{1}{2} \bigl(p_v[k_j - \mathrm{cut}(S_j;G)] + p_v[k_j + \mathrm{cut}(S_j;G)] \bigr) \nonumber\\
	& \leq \alpha + \frac{1}{2} \bigl(p_v[k_j - \Phi(S_j;G) \wb{k}_j] + p_v[k_j + \Phi(S_j;G) \wb{k}_j]  \bigr) \nonumber \\
	& \leq \alpha + \frac{1}{2} \bigl(p_v[k_j - \phi \wb{k}_j] + p_v[k_j + \phi \wb{k}_j]\bigr) \nonumber
	\end{align}
	and subtracting $k_j/2m$ from both sides, we get
	\begin{equation}
	\label{eqn:mixing_time_PPR_pf1}
	h^{(\alpha)}(k_j) \leq \alpha + \frac{1}{2} \bigl(h^{(\alpha)}(k_j - \phi \wb{k}_j) + h^{(\alpha)}(k_j +  \phi \wb{k}_j) \bigr)
	\end{equation}
	From this point, we divide our analysis into cases.

	\textbf{Case 1.}
	Assume $k_j - 2 \phi \wb{k}_j$ and $k_j + 2 \phi \wb{k}_j$ are both in $[K_0,2m  - K_0]$. We are therefore in a position to apply our inductive hypothesis to \eqref{eqn:mixing_time_PPR_pf1}, yielding
	\begin{align*}
	h^{(\alpha)}(k_j) & \leq \alpha + \alpha(t-1) \frac{1}{2}\biggl(L_{K_0}(k_j - \phi \wb{k}_j) + L_{K_0}(k_j + \phi \wb{k}_j) + C(K_0;h^{(\alpha)})\bigl(\sqrt{\wb{k_j - \phi \wb{k}_j}} + \sqrt{\wb{k_j + \phi \wb{k}_j}}\bigr)\left(1 - \frac{\phi^2}{8}\right)^{t-1} \biggr) \\
	& \leq \alpha t + L_{K_0}(k;h^{(\alpha)}) + \frac{1}{2}\biggl(C(K_0;h^{(\alpha)})\bigl(\sqrt{\wb{k_j - \phi \wb{k}_j}} + \sqrt{\wb{k_j + \phi \wb{k}_j}}\bigr)\left(1 - \frac{\phi^2}{8}\right)^{t-1} \biggr) \\
	& \leq \alpha t + L_{K_0}(k;h^{(\alpha)}) + \frac{1}{2}\biggl(C(K_0;h^{(\alpha)})\bigl(\sqrt{\wb{k}_j - \phi \wb{k}_j} + \sqrt{\wb{k}_j + \phi \wb{k}_j}\bigr)\left(1 - \frac{\phi^2}{8}\right)^{t-1} \biggr).
	\end{align*}
	and therefore
	\begin{equation*}
	h^{(\alpha)}(k_j) \leq  \alpha t + L_{K_0}(k;h^{(\alpha)}) + \frac{C(K_0;h^{(\alpha)})}{2}\cdot \sqrt{\wb{k}_j}\cdot\left(2 - \frac{\phi^2}{4}\right)\left(1 - \frac{\phi^2}{8}\right)^{t-1} = \alpha t + L_{K_0}(k;h^{(\alpha)}) + C(K_0;h^{(\alpha)})\sqrt{\wb{k}_j}\left(1 - \frac{\phi^2}{8}\right)^{t}.
	\end{equation*}

	\textbf{Case 2.} Otherwise one of $k_j - 2 \phi \wb{k}_j$ or $k_j + 2 \phi \wb{k}_j$ is not in $[K_0,2m  - K_0]$. Without loss of generality assume $k_j < m$, so that (i) we have $k_j - 2 \phi \wb{k}_j < K_0$ and (ii) $k_j + (k_j - K_0) \leq 2m - K_0$. By the concavity of $h$, and applying the inductive hypothesis to $h^{(\alpha)}2k_j - K_0)$, we have
	\begin{align*}
	h^{(\alpha)}(k_j) & \leq \alpha + \frac{1}{2}\Bigl(h^{(\alpha)}(K_0) + h\bigl(k_j + (k_j - K_0)\bigr)\Bigr) \\
	& \leq\alpha + \frac{\alpha(t - 1)}{2} + \frac{1}{2}\Bigl(L_{K_0}(K_0;p^{\alpha}) + L_{K_0}(2k_j - K_0\bigr) + C(K_0;h^{(\alpha)})\sqrt{\wb{2k_j - K_0}}\left(1 - \frac{\phi^2}{8}\right)^{t - 1}\Bigr) \\
	& \leq \alpha t + L_{K_0}(k_j) + C(K_0;h^{(\alpha)}) \frac{\sqrt{2\wb{k}_j}}{2} \left(1 - \frac{\phi^2}{8}\right)^{t - 1} \\
	& \leq \alpha t + L_{K_0}(k_j) + C(K_0;h^{(\alpha)}) \sqrt{\wb{k}_j} \cdot \left(1 - \frac{\phi^2}{8}\right)^{t}
	\end{align*}
\end{proof}

@@

\section{Sample-to-population bounds}
\label{sec:sample_to_population}
In this section, we prove Propositions~\ref{prop:sample_to_population_1} and~\ref{prop:sample_to_population_2}, by establishing high-probability finite-sample bounds on various functionals of the random graph $G_{n,r}$: cut, volume, and normalized cut~(\ref{subsec:sample_to_population_ncut}), minimum and maximum degree, and local spread~(\ref{subsec:sample_to_population_local_spread}), and conductance~(\ref{subsec:sample_to_population_conductance}). To establish these results, we will use several different concentration inequalities, and we begin by reviewing these in~(\ref{subsec:concentration}). Throughout, we denote the empirical probability of a set $\mc{S} \subseteq \Rd$ as $\Pbb_n(\mc{S}) = \sum_{i = 1}^{n} \1\{x_i \in \mc{S}\}/n$, and the conditional (on being in $\mc{C}$) empirical probability as $\wt{\Pbb}_n = \sum_{i = 1}^{n} \1\{x_i \in (\mc{S} \cap \mc{C})\}/\wt{n}$, where $\wt{n} = |\mc{C}[X]|$ is the number of sample points that are in $\mc{C}$. For a probability measure $\Qbb$, we also write
\begin{equation}
\label{eqn:population_minmax_degree}
d_{\min}(\Qbb) := \inf_{x \in \mathrm{supp}(\mbb{Q})} \deg_{\Pbb,r}(x),~~\textrm{and}~~d_{\max}(\Qbb) := \sup_{x \in \mathrm{supp}(\mbb{Q})} \deg_{\Pbb,r}(x).
\end{equation}

@@

Many graph functionals are order-2 U-statistics, and we use Hoeffding's inequality to control the deviations of these functionals from their expectations. Recall that $U_n$ is an order-2 U-statistic with kernel $\varphi: \Rd \times \Rd \to \Reals$ if 
\begin{equation*}
U_n = \frac{1}{n(n-1)}\sum_{i = 1}^{n} \sum_{j \neq i} \varphi(x_i,x_j).
\end{equation*}
We write $\norm{\varphi}_{\infty} = \sup_{x,y} |\varphi(x,y)|$. 
\begin{lemma}[Hoeffding's Inequality for Order-2 U-statistics.]
	\label{lem:hoeffding}
	Fix $\delta \in (0,1)$. Assume $\norm{\varphi}_{\infty} \leq 1$. Then,
	\begin{equation*}
	(1 - \delta) \mathbb{E}[U_n] \leq U_n \leq (1 + \delta) \mathbb{E}[U_n]
	\end{equation*}
	with probability at least $1 - 2 \exp\left(-\delta^2 (\mathbb{E}[U_n])^2 n\right)$.
\end{lemma}

@@

	For any $x \in \Rd$, the remainder terms can be upper bounded: since $\Delta_{12} \geq 0$,

@@

	if $0 \leq \Delta_{12} \leq r/(4(d - 1))$,

@@

	To derive~\eqref{pf:conductance_lb_transportation_distance_3}, suppose that $\mc{R} \subseteq \mathrm{supp}(\mbb{Q}_2)$ and $\mc{R} \not\in \mathfrak{L}(\mbb{Q}_1,\mbb{Q}_2)$. Without loss of generality, we shall assume that $\vol_{\mbb{Q}_2,r}(\mc{R}) \leq (1 - \Psi_r(\mbb{Q}_1)) \cdot d_{\min}(\mbb{Q}_2)^2$ (otherwise we can work with respect to $\mc{R}^c$.)  Then, for all $x \in \mc{R}$,

@@

\section{Additional results: aPPR and Consistency of PPR}
\label{sec:appr_misclassification_error}

\subsection{Consistency}
\label{subsec:consistent_recovery_density_clusters}

The symmetric set difference does not measure whether \smash{$\wh{C}$}
can (perfectly) distinguish any two distinct clusters \smash{$\mc{C}_{\lambda},\mc{C}_{\lambda}' \in 
	\mathbb{C}_f(\lambda)$}. We therefore also study a second notion of cluster 
estimation, first introduced by \citet{hartigan1981}, and defined
asymptotically.
